---
title: 'Lab 3: Exploring Happiness'
subtitle: "Due: Thursday 10/03/2019 by 10 pm"
author: "Ashley Anderson & Hexiang Zhang"

date: "09/26/2019"
output: html_document
---


## Background

The General Social Survey (GSS) has been run by NORC every other year since 1972 to keep track of current opinions across the United States. We are working with an excerpt of the dataset that covers years 1972 to 2016. The variable names along with a brief description are as follows: 

| Variable      | Description   |    
| ------------- |:-------------:| 
| Year          | Year for this respondent | 
| Age           | Age of respondent | 
| Sex           | Sex of respondent     | 
|....           | .....      | 

Now it's important to understand what each column stands for. You can find a codebook with explanations for each of the variables at https://gssdataexplorer.norc.org/


```{r}
library(classdata)
library(dplyr)

happy <- readRDS("HAPPY.rds")
names(happy) = tolower(names(happy))

```

## Data Cleaning 

Initial exploration revealed that some data processing was needed before we could proceed with further visualization. The first table below shows the number of missing values for all variables before data cleaning, labeled as "IAP", "DK", and "NA". After this discovery, we changed all missing values labeled as "DK" and "IAP" and relabeled them as "NA". The second table belows shows this was successful. 
```{r}
na.table <- data.frame()
for(i in 1:ncol(happy)){
  na.table[1,i] <- sum(is.na(happy[i]))
  na.table[2,i] <- length(grep("IAP", happy[,i]))
  na.table[3,i] <- length(grep("DK", happy[,i]))
}
names(na.table) <- colnames(happy)
rownames(na.table) <- c("NAs","IAP","DK")
na.table

happy2 <- happy

na_codes <- c("IAP", "DK", "")
for (i in seq_along(happy2)) {
  happy2[[i]][happy2[[i]] %in% na_codes] <- NA
}

na.table2 <- data.frame()
for(i in 1:ncol(happy2)){
  na.table2[1,i] <- sum(is.na(happy2[i]))
  na.table2[2,i] <- length(grep("IAP", happy2[,i]))
  na.table2[3,i] <- length(grep("DK", happy2[,i]))
}
names(na.table2) <- colnames(happy2)
rownames(na.table2) <- c("NAs","IAP","DK")
na.table2
```


As we can see below, a lot of variables were also cast into the wrong variable type. Age appears to be a character variable, and should be numeric. All categorical variables needed to be changed to type factor (except partyid and polyviews). There were 335 instances where age appeared as "89 OR OLDER". We changed this for all instances that were labeled as "89 OR OLDER" and labeled them as 89 so the variable age could be changed to type numeric. We can now see below that all variabels are now cast into the correct variable type. 

```{r}
str(happy)
length(grep("89 OR OLDER", happy2$age))

for (i in 1:nrow(happy2)) {
  age <- happy2[i, 'age']
  if(!is.na(age)){
    if(age=="89 OR OLDER") {
      happy2[i, 'age'] <- 89
    }
  }
}

happy2$happy <- as.factor(happy2$happy)
happy2$age <- as.numeric(happy2$age)
happy2$sex <- as.factor(happy2$sex)
happy2$marital <- as.factor(happy2$marital)
happy2$degree <- as.factor(happy2$degree)
happy2$finrela <- as.factor(happy2$finrela)
happy2$health <- as.factor(happy2$health)

str(happy2)
```

Factors were also organized into meaningful order as follows: We ordered marital status by age and health by age.
```{r}
marital.df <- as.data.frame(aggregate(happy2$age, by=list(happy2$marital), mean, na.rm=TRUE))
marital.order <- c(order(marital.df[,"x"]))
happy2$marital <- factor(happy2$marital,levels(happy2$marital)[marital.order])

health.df <- as.data.frame(aggregate(happy2$age, by=list(happy2$health), mean, na.rm=TRUE))
health.order <- c(order(health.df[,"x"]))
happy2$health <- factor(happy2$health,levels(happy2$health)[health.order])

str(happy2)
```


## Data Exploration 

Two variables of interest are `...` and `...`. Their distributions are as follow: 
```{r}
#insert relevant R code here
#include graphics. DESCRIBE your plot in context. Make sure it includes axis labels and appropriate title. 
```

The distribution for `...` is `....`, suggesting that `....` 




